### 2016-05-14 Paper/Paper.tex
Need abstract

\section{Introduction}
\subsection{Carbon Storage}
Scope to US bc available data

Motivation and whats what

\subsection{Potential Storage Site: Saline Formations}

\begin{wrapfigure}{r}{0.5\textwidth} %this figure will be at the right
    \centering
    \includegraphics[width=0.5\textwidth]{saline_tight}
    \caption{\label{salinetight} A map of the United States overlaid with geological saline resources for potential carbon storage (blue), stationary sources of petroleum and natural gas (brown) and other oil and gas resources (red) \cite{atlas}.}
\end{wrapfigure}

\par Of possible candidates for long term carbon sequestration, saline formations have emerged as a promising candidate. Such formations are prevalent throughout North America. These brine coated layers of permeable rock are able to store carbon my way of "solubility trapping, mineral trapping, structural trapping and residual trapping."

The more than 180,000 potential injection sites in the United States have been estimated by the U.S. Department of Energy to have capacity of over 12 trillion tons of CO$_2$. Saline formations are promising also because of their proximity to CO$_2$ point sources, allowing easy transition of U.S. energy assets to "near-zero carbon emissions via low-cost carbon storage retrofits." \cite{whysaline} Furthermore, there is much existing technology and regulatory acceptance with respect to injections into saline reservoirs as brines are frequently injected into saline reservoirs in EOR and the U.S. EPA has designated some deep saline formations for hazardous waste disposal.

\par There are currently at least 3 saline aquifer injection projects that have been undertaken in the continental United States, which have together injected over a million tons of CO$_2$ to date. Each of these projects operates on a time scale of at least twenty years, allocating at least 10 years to "characterizing geologic and terrestrial opportunities for carbon storage and identifying CO$_2$ stationary sources within the territories of the individual RCSPs and evaluating promising CO$_2$ storage opportunities through a series of small-scale field projects"  \cite{midwestinject}, \cite{midwestinject2}, \cite{southeastinject}.

Important Charecterists

\subsubsection{Associated Risk} 
Maybe use that earthquake graph here....

\section{Methodology}
\subsection{Overview}
As described in the previous section, saline aquifers are luckily a far vast enough resource for us to currently store the desired amount of carbon. However, selecting the best aquifers for storage is a tremendous task considering the quantity of potential aquifers ($\approx180,000$) and the factors that characterize each one. For this reason we seek to develop a tool to immediately help identify some of the best candidate deep saline aquifers in the United States. 
The fundamental workings of this tool will be in the characterization of potential aquifers based on a series of characteristics as per section \ref{wellchar}. Having fully characterized each candidate aquifer, we fully employ a Fuzzy c-Means clustering analysis to each of the $\approx180,000$, 12 dimensional vectors representing the system to create a model that will determine how strongly each aquifer is characterized by each of the 12 features, to be discusses further in section \ref{fuzz}. Once this model is created, it will allow researchers to essentially filter potential storage sites by different characteristics and determine the relative similarity of unique sets of sites. For example, we will be able to observe all aquifers that have at least a given storage capacity and/or are at least a given distance from a major fault line.

\noindent \emph{N.B. For the code that produced the results of this study, please see the repository \\at} https://github.com/joshuazweig/Carbon-Storage-Reservoir-Ranker \emph{where you can download the code and the data and run it yourself}

@@

\begin{figure}[H]%this figure will be at the right
  \centering
  \includegraphics[width=0.7\textwidth]{induced}
   \caption{\label{inducedfig} From \cite{seis_induced}. Count of M $\geq3$ earthquakes in the central and eastern United States from 1973 to April 2015. Two abrupt increases in the earthquake rate occurred in 2009 and 2013. Red dots represent earthquakes that occurred between 2009 and April 2015, and blue dots represent earthquakes that occurred between 1973 and 2008. Prior to 2009, earthquakes were spread across the United States. Beginning in 2009 the earthquakes are tightly clustered in a few areas.}
\end{figure}

@@

\begin{figure}[H]
\centering
\caption{\label{clusterfig} Each Figure (a - e) shows the clusters resulting from the running of the FPC algorithm for unique values if $k$. These results actually occur in 9-dimensional space, so a 6 degree Principal Component Analysis was used to reduce the dimensionality of each vector to 3 so that they could be pictured in these figures. For this reason, some of the detail gets lost, but the clusters are all still discernable. In Section \ref{danda} we will find (d), where $k = 9$, to be the preferred clustering analysis, hence it is colored differently than the rest of the plots.}
\begin{tabular}{cc}
  \includegraphics[width=65mm]{k5} &   \includegraphics[width=65mm]{k6} \\
(a) $k = 5, FPC = 0.6667$ & (b) $k = 6, FPC = .8333$ \\[6pt]
 \includegraphics[width=65mm]{k7} &   \includegraphics[width=65mm]{k8} \\
(c) $k = 7, FPC = .8333$ & (d) $k = 8, FPC =  0.6667$ \\[6pt]
\end{tabular}
\end{figure}
\begin{figure}[H]
\centering
\begin{tabular}{cc}
 \includegraphics[width=65mm]{k9} &   \includegraphics[width=65mm]{k10} \\
(d) $k = 9, FPC = .6111$ & (e) $k = 10, FPC = 1$ \\[6pt]
\end{tabular}
\end{figure}

@@

\par Upon inspection, it is difficult to isolate exactly the dimensions and thresholds that gain a site entry into cluster $A$. However, the dimensions that are prevalent in $B$ and $C$ allow is to form something of a negative definition of $A$. 
(You should discuss this more.......)

@@

\subsection{Better Characterization of Saline Formations}\label{betterchar}
The 12 features isolated for characterization are a minimal set for reasonable characterization of saline formations. Better and more detailed (higher dimensional) characterizations of these formations would necessarily allow for more data for Fuzzy c-Means clustering to base its similarity measures on, thereby increasing the accuracy and effectiveness of this tool. Further dimensions include the existence of a caprock over the layer of porous rock. The existence of such a caprock is extremely important to the viability of a potential site as it plays an important part in keeping the injected carbon below the surface and in place. I would like to further investigate and apply some of the findings of \cite{natcarb_risk} and \cite{PredRisk2013} in order to represent every formation by the most relevantly detailed dimensions possible.

@@

\section{Conclusion}

Value judgements -- who is to say how imporatant it is to not do this under a national park, etc- --- this system allows for quick hypotheses testing based on different criteria, on different values and best of all can isolate some that meet everyones turf the best, givingbetter information of better options

@@

@@

\bibitem{PredRisk2013}
Balashov, Victor N., et al.
\textit{Predictive modeling of CO 2 sequestration in deep saline sandstone reservoirs: impacts of geochemical kinetics}
Applied geochemistry 30 (2013): 41-56.

>>>>

\section{Introduction}
\subsection{Carbon Storage}
Rising levels of CO$_2$ in Earth's atmosphere have generated wide spread concern about the global ecology and climate. For the first time in recent history, levels of CO$_2$ in Earth's atmosphere have risen above 400ppm, not much shy of what experts say could be a catastrophic level at 450ppm. Contemporarily, with the global attention received to this issue by COP 21 and other forums of politicians and scientists, this is an issue whose importance seems to need little motivation.
Following from this, the world must move towards a carbon neutral energy landscape. This does not necessarily imply zero carbon dioxide output, but will very likely result in a world where we capture and store or utilize as much carbon as we put out. To achieve a world in which we do not have to remove some of the carbon from the atmosphere, we would have already have had to make large steps towards reducing our $85\%$ fossil fuel based energy infrastructure to a zero carbon output one. We have not made steps in nearly large enough magnitude to avoid having to take some of the carbon out of the atmosphere at some point. 
This has motivated a world wide effort of Carbon Capture, Utilization and Storage (CCUS). Storage can take many different forms. Possible options include ocean disposal, one of the early suggestions to the storage problem. Since the oceans and atmosphere work towards a carbon balance, this suggestion takes the form of accelerating that process. There are many ecological risks that this imposes and is likely difficult to achieve the international widespread public acceptance necessary. A solution increasing in popularity is mineral sequestration. Leveraging abundant minerals including olivine and serpentines to produce carbonates is potentially an extremely viable solution, but we do not currently have the technology and/or economic incentives aligned to apply this technique at scale right now. A third and potentially attractive option is geological disposal. In this model of sequestration, supercritical CO$_2$ is injected at great depths into different geological resources, applying many of the well developed injection techniques of the Oil and Gas industry as in the case of enhanced oil recovery. There are currently many projects underway that are injecting CO$_2$ underground, meriting that this approach be well studied.

\subsection{Potential Storage Resource: Saline Formations}
\setlength{\intextsep}{2pt}%
\begin{wrapfigure}{L}{0.4\textwidth} %this figure will be at the right
    \centering
    \includegraphics[width=0.4\textwidth]{sdiag}
    \caption{\label{sdiagfig}  From \cite{atlas}. A diagram showing the typical workflow of capturing and injecting CO$_2$ into underground saline formations. }
\end{wrapfigure}

\par Of possible candidates for long term carbon sequestration, saline formations have emerged as a promising candidate. Such formations are prevalent throughout North America. These brine coated layers of permeable rock are able to store carbon my way of "solubility trapping, mineral trapping, structural trapping and residual trapping." Figure \ref{sdiagfig} shows the typical process by which captured carbon dioxide can be injected into saline formations producing brine as a byproduct \cite{atlas}.

The more than 180,000 potential injection sites in the United States have been estimated by the U.S. Department of Energy to have capacity of over 12 trillion tons of CO$_2$. Saline formations are promising also because of their proximity to CO$_2$ point sources, allowing easy transition of U.S. energy assets to "near-zero carbon emissions via low-cost carbon storage retrofits." \cite{whysaline} Furthermore, there is much existing technology and regulatory acceptance with respect to injections into saline reservoirs as brines are frequently injected into saline reservoirs in EOR and the U.S. EPA has designated some deep saline formations for hazardous waste disposal. There are many signs point to saline formations as part of the carbon storage solution. 
\pagebreak
\setlength{\intextsep}{2pt}%
\begin{wrapfigure}{R}{0.4\textwidth} %this figure will be at the right
    \centering
    \includegraphics[width=0.4\textwidth]{saline_tight}
    \caption{\label{salinetight} A map of the United States overlaid with geological saline resources for potential carbon storage (blue), stationary sources of petroleum and natural gas (brown) and other oil and gas resources (red) \cite{atlas}.}
\end{wrapfigure}

\par There are currently at least 3 saline aquifer injection projects that have been undertaken in the continental United States, which have together injected over a million tons of CO$_2$ to date. Each of these projects operates on a time scale of at least twenty years, allocating at least 10 years to "characterizing geologic and terrestrial opportunities for carbon storage and identifying CO$_2$ stationary sources within the territories of the individual RCSPs and evaluating promising CO$_2$ storage opportunities through a series of small-scale field projects"  \cite{midwestinject}, \cite{midwestinject2}, \cite{southeastinject}. Some of the important aspects of these endeavors of characterization, which are very important to this work, will be discussed in subsection \ref{wellchar}. 
\setlength{\intextsep}{2pt}%
\begin{wrapfigure}{l}{.4\textwidth}%this figure will be at the right
  \centering
  \includegraphics[width=0.4\textwidth]{induced}
   \caption{\label{inducedfig} From \cite{seis_induced}. Count of M $\geq3$ earthquakes in the central and eastern United States from 1973 to April 2015. Two abrupt increases in the earthquake rate occurred in 2009 and 2013, with an increase in EOR activity. Red dots represent earthquakes that occurred between 2009 and April 2015, and blue dots represent earthquakes that occurred between 1973 and 2008.}
\end{wrapfigure}
\subsubsection{Potential Risks and Drawbacks of Saline Formations} 
One major potential risk of large scale injections of supercritical carbon is that of induced seismicity. Most induced earthquaked and seismic activity in the U.S. are caused by the deep injection of fluids, such as would be done in saline formation sequestration \cite{induced}. This is certainly an important consideration and was well studied by the characterization projects in \cite{midwestinject}, \cite{midwestinject2}, \cite{southeastinject}. In addition to public safety directly from earthquakes, any sort of induced seismic activity is potentially detrimental as it can open passageways for CO$_2$ to flow deep underground, increasing the loss rate and difficulty of the monitoring and verification aspect of carbon storage. If there is high risk of loosing the carbon once we put it underground, it is unlikely we will want to do it in the first place.

An additional issue with storage, especially in saline formations, is the massive injection volumes. In the last section, the volumes that would need to be injected and number of potential sites investigates and managed for this storage is almost intractable. However, the NATCARB project is making great strides along these lines and this work seeks to mitigate the risk involved with both factors outlined here.

\section{Methodology}
\subsection{Overview}
As described in the previous section, saline aquifers are luckily a far vast enough resource for us to currently store the desired amount of carbon. However, selecting the best aquifers for storage is a tremendous task considering the quantity of potential aquifers ($\approx180,000$) and the factors that characterize each one. For this reason we seek to develop a tool to immediately help identify some of the best candidate deep saline aquifers in the United States. 
The fundamental workings of this tool will be in the characterization of potential aquifers based on a series of characteristics as per section \ref{wellchar}. Having fully characterized each candidate aquifer, we fully employ a Fuzzy c-Means clustering analysis to each of the $\approx180,000$, 12 dimensional vectors representing the system to create a model that will determine how strongly each aquifer is characterized by each of the 12 features, to be discusses further in section \ref{fuzz}. Once this model is created, it will allow researchers to essentially filter potential storage sites by different characteristics and determine the relative similarity of unique sets of sites. For example, we will be able to observe all aquifers that have at least a given storage capacity and/or are at least a given distance from a major fault line.  
\\
\noindent \emph{N.B. For the code that produced the results of this study, please see the repository \\at} https://github.com/joshuazweig/Carbon-Storage-Reservoir-Ranker \emph{where you can download the code and the data and run it yourself}

@@

@@

\begin{figure}[H]
\centering
\caption{\label{clusterfig} Each Figure (a - e) shows the clusters resulting from the running of the FPC algorithm for unique values if $k$. These results actually occur in 9-dimensional space, so a 6 degree Principal Component Analysis was used to reduce the dimensionality of each vector to 3 so that they could be pictured in these figures. For this reason, some of the detail gets lost, but the clusters are all still discernable. In Section \ref{danda} we will find (d), where $k = 9$, to be the preferred clustering analysis, hence it is colored differently than the rest of the plots.}
\begin{tabular}{cc}
  \includegraphics[width=65mm]{k5} &   \includegraphics[width=65mm]{k6} \\
(a) $k = 5, FPC = 0.6667$ & (b) $k = 6, FPC = .8333$ \\[6pt]
\end{tabular}
\end{figure}
\begin{figure}[H]
\centering
\begin{tabular}{cc}
 \includegraphics[width=65mm]{k7} &   \includegraphics[width=65mm]{k8} \\
(c) $k = 7, FPC = .8333$ & (d) $k = 8, FPC =  0.6667$ \\[6pt]
 \includegraphics[width=65mm]{k9} &   \includegraphics[width=65mm]{k10} \\
(d) $k = 9, FPC = .6111$ & (e) $k = 10, FPC = 1$ \\[6pt]
\end{tabular}
\end{figure}

@@

\par Upon inspection, it is difficult to isolate exactly the dimensions and thresholds that gain a site entry into cluster $A$. However, the dimensions that are prevalent in $B$ and $C$ allow is to form something of a negative definition of $A$. Analysis reveals that cluster $B$ contains $83.3\$$ of wells in the set of wells with the highest $50\%$ of scores. This seems to be the overwhelming characteristic uniting formation vectors in $B$. From this we can conjecture that the majority of the vectors in $A$ scored in at most the 50th percentile in terms of their proximity to active fault lines. The elements of $C$ are significantly more difficult to decipher. It is expected that any machine learning algorithm finds patterns not immediately noticeable by humans, and further analysis will be necessary to determine the pattern that $C$ seems to have isolated.

@@

\subsection{Better Characterization of Saline Formations}\label{betterchar}
The 12 features isolated for characterization are a minimal set for reasonable characterization of saline formations. Better and more detailed (higher dimensional) characterizations of these formations would necessarily allow for more data for Fuzzy c-Means clustering to base its similarity measures on, thereby increasing the accuracy and effectiveness of this tool. Further dimensions include the existence of a caprock over the layer of porous rock. The existence of such a caprock is extremely important to the viability of a potential site as it plays an important part in keeping the injected carbon below the surface and in place. I would like to further investigate and apply some of the findings of \cite{atlas}, \cite{natcarb_risk} and \cite{PredRisk2013} in order to represent every formation by the most relevantly detailed dimensions possible.

@@

\section{Conclusion}
\par In this paper, I have devised a new approach to beginning the process of saline formation characterization. Previously, with so many potential sites and so many unique characterizations of each, it would have been impossible to take a comprehensive look at every possible candidate formation, resulting in some of the best options likely being glossed over. Through the fuzzy clustering analysis presented in this paper I have already given a reasonable place for researchers to focus their search for formations in which to begin carbon injections, namely cluster $A$.

\par However, there remains a use case just as important for what I have created here and that is the modularity of the way in which extra-geological dimensions are scored in creating the vector representation of the formations. This is tremendously important and gives the analysis its power because it lets researchers investigate the saline formation landscape based on their own value judgements and thresholds of what is important and what is not. This really hits at a major underlaying issue in the tackling of our carbon problem, value judgements. Who is to say how important it is if there is an earthquake in a national park resulting from induced seismicity? Who is to say the real cost of another 20ppm of CO$_2$ in the atmosphere. Everyone makes their own, unique value judgements on the climate issues were facing today, which is one of the reasons they feel so insurmountable and that it is difficult to achieve real consensus and real action. Rather than that being a weakness, the system I have devised here embraces it at its core in easily allowing anyone interested to weight different factors as they wish to approximate their own values and their own use case.

@@

\bibitem{salinestore}
Nordbotten, J.M., Celia, M.A., Bachu, S.
\textit{Injection and Storage of CO$_2$ in Deep Saline Aquifers: Analytical Solution for CO$_2$ Plume Evolution During Injection}
Transport in Porous media 58.3 (2005): 339-360.

@@

\bibitem{PredRisk2013}
Balashov, Victor N., et al.
\textit{Predictive modeling of CO$_2$ sequestration in deep saline sandstone reservoirs: impacts of geochemical kinetics}
Applied geochemistry 30 (2013): 41-56.
